An investigation into the contentious phylogenetic results of the *Mixta* species

### 2021-04-23 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

---
title: "An investigation into the contentious phylogenetic results of the *Mixta* species"
output:
  word_document:
    reference_docx: Style_Manuscript.docx
    fig_caption: true
    # toc: true
    # toc_depth: 5
bibliography: References.bib
csl: IJSEM.csl
---
```{r Packages, echo = FALSE, message = FALSE, warning=FALSE}
library("knitr")
library("tidyverse")
library("ggplot2")
library("ggpubr")
library("RColorBrewer")
library("extrafont")
theme_set(theme_classic())
```

@@

## Spatial Patterns from Average Site Identity

@@

# Old Results
<!-- ```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 6. The most similar taxa to *M. calida* genes for nucleotide sequences around its genome. The most similar taxon is that which sequence is most identical to the *M. calida* sequence. Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5} -->
<!-- FI_calida_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6_CircularIdentities_NT.csv", -->
<!--                          stringsAsFactors = FALSE) %>% -->
<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
<!--                                               "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE)) -->

<!-- ggplot(FI_calida_NT, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
<!--   coord_polar() + -->
<!--   scale_fill_brewer(palette = "Paired", -->
<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
<!--                                "P. syringae")) + -->
<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
<!--   labs(fill = "Most Similar Taxa") + -->
<!--   theme_bw() + -->
<!--   theme(legend.text = element_text(face = "italic"), -->
<!--         axis.title = element_blank(), -->
<!--         axis.text.y = element_blank(), -->
<!--         axis.ticks.y = element_blank(), -->
<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
<!-- ``` -->

<!-- ```{r Figure7, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 7. The most similar taxa to *M. calida* genes for amino acid sequences around its genome. The most similar taxon is that which sequence is most identical to the *M. calida* sequence. Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5} -->
<!-- FI_calida_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure7_CircularIdentities_AA.csv", -->
<!--                          stringsAsFactors = FALSE) %>% -->
<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
<!--                                               "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE)) -->

<!-- ggplot(FI_calida_AA, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
<!--   coord_polar() + -->
<!--   scale_fill_brewer(palette = "Paired", -->
<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
<!--                                "P. syringae")) + -->
<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
<!--   labs(fill = "Most Similar Taxa") + -->
<!--   theme_bw() + -->
<!--   theme(legend.text = element_text(face = "italic"), -->
<!--         axis.title = element_blank(), -->
<!--         axis.text.y = element_blank(), -->
<!--         axis.ticks.y = element_blank(), -->
<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
<!-- ``` -->

<!-- ```{r Table5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE} -->
<!-- stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE) -->

<!-- same_spp <- data.frame(table(stats_on_all$Same_Species)) -->
<!-- same_spp <- same_spp[which(same_spp$Var1 != "Nope"), ] -->

<!-- same_spp <- rbind(same_spp, data.frame(Var1 = "T_ptyseos", Freq = 0)) -->

<!-- colnames(same_spp) <- c("Species", "Number of Genes") -->

<!-- same_spp <- mutate(same_spp, -->
<!--                    Species = case_when(Species == "E_amylovora" ~ "E. amylovora", -->
<!--                                        Species == "E_tasmaniensis" ~ "E. tasmaniensis", -->
<!--                                        Species == "P_agglomerans" ~ "P. agglomerans", -->
<!--                                        Species == "P_septica" ~ "P. septica", -->
<!--                                        Species == "T_ptyseos" ~ "T. ptyseos", -->
<!--                                        Species == "T_saanichensis" ~ "T. saanichensis", -->
<!--                                        Species == "E_cloacae" ~ "E. cloacae")) %>% -->
<!--   mutate(Species = factor(Species, levels = c("P. agglomerans", "P. septica", "E. amylovora", "E. tasmaniensis", -->
<!--                                               "T. ptyseos", "T. saanichensis", "E. cloacae"), ordered = TRUE)) -->

<!-- same_spp <- arrange(same_spp, Species) -->

<!-- kable(same_spp, caption = "Table 5. The number of genes that had each species as the closest or most similar relative for all four analyses.") -->

<!-- # table(stats_on_all$Same_Genus) -->
<!-- # table(stats_on_all$NT_Same) -->
<!-- # table(stats_on_all$AA_Same) -->
<!-- # table(stats_on_all$Rel_Same) -->
<!-- # table(stats_on_all$Ide_Same) -->
<!-- ``` -->

<!-- ```{r Figure8, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 8. Sample for first 100 genes.", fig.width=6.5, fig.height=3} -->
<!-- all_results_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureN_AllResultsCalida.csv", -->
<!--                                stringsAsFactors = TRUE) %>% -->
<!--   pivot_longer(cols = c(Rel_NT:Ide_AA), names_to = "Analysis", values_to = "Closest_Similar") %>% -->
<!--   mutate(Closest_Similar = factor(Closest_Similar, -->
<!--                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", -->
<!--                                              "E_cloacae", "P_syringiae"), -->
<!--                                   ordered = TRUE), -->
<!--          Analysis = factor(Analysis, levels = c("Ide_AA", "Rel_AA", "Ide_NT", "Rel_NT"), ordered = TRUE), -->
<!--          Gene_Name = reorder(Gene_Name, ID)) -->

<!-- ggplot(all_results_calida[1:400, ], aes(x = Gene_Name, y = Analysis, fill = Closest_Similar)) + -->
<!--   geom_tile() + -->
<!--   scale_fill_brewer(palette = "Paired", -->
<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
<!--                                "P. syringae")) + -->
<!--   scale_y_discrete(labels = c("AA Identity", "AA Closest", "NT Identity", "NT Closest")) + -->
<!--   labs(fill = "Closest Relative or \n Most Similar Taxa") + -->
<!--   theme(legend.text = element_text(face = "italic"), -->
<!--         # legend.position = "bottom", -->
<!--         text = element_text(size = 12,  family = "Times New Roman"), -->
<!--         axis.text.x = element_text(angle = 45, hjust = 1))               # Rotates x axis labels -->

<!-- stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE) -->

<!-- # table(stats_on_all$Same_Species) -->
<!-- # table(stats_on_all$Same_Genus) -->
<!-- # table(stats_on_all$NT_Same) -->
<!-- # table(stats_on_all$AA_Same) -->
<!-- # table(stats_on_all$Rel_Same) -->
<!-- # table(stats_on_all$Ide_Same) -->
<!-- ``` -->

<!-- ```{r Figure9, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 9. Maximum likelihood gene trees for ttuB for nucleotide (A) and amino acid sequences (B).", fig.width=3.25} -->
<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_NT.png") -->
<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_AA.png") -->
<!-- ``` -->

<!-- ```{r Figure10, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 10. Maximum likelihood gene trees for secY for nucleotide (A) and amino acid sequences (B). Since bifucations could not be seen in radial trees, condensed trees were computed with a 70% cutoff for both the nucleotide sequnces (C) and amino acid sequences (D).", fig.width=3.25} -->
<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_NT.png") -->
<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_AA.png") -->
<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_NT70.png") -->
<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_AA70.png") -->
<!-- ``` -->

>>>>

---
title: "An investigation into the contentious phylogenetic results of the *Mixta* species"
output:
  word_document:
    reference_docx: Style_Manuscript.docx
    fig_caption: true
    # toc: true
    # toc_depth: 5
bibliography: References.bib
csl: IJSEM.csl
---
```{r Packages, echo = FALSE, message = FALSE, warning=FALSE}
library("knitr")
library("tidyverse")
library("ggplot2")
library("ggpubr")
library("RColorBrewer")
library("extrafont")
theme_set(theme_classic())

Spp <- c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", 
         "E_cloacae", "P_syringae")
```

@@

## Spatial Patterns from Average Site Identity
Positional effects were investigated in the site identity results. Only the results of *M. calida* are shown because the *M. gaviniae* results showed similar patterns. The genes most similar to either of the *Pantoea* species and *E. cloacae* were evenly dispersed across the *Mixta* chromosome for both the NTS and AAS (Figure 5A and 5B). Similar to the genetic distance results, genes that are labelled as being closely related to either of the *Tatumella* species are located only on right side of chromosome, and genes that are closely related to *E. amylovora* are predominantly on the right side as well (Figure 5A). Unexpectedly, the lower left quadrant of the chromosome contains primarily genes that are closely related to either of the *Pantoea* species with only a few exceptions (Figure 5A). 	In the AAS analysis, genes closely related to any of the other taxa were more evenly distributed across the genome (Figure 5B).

@@

## A Comparison of the Four Analyses
The four above analyses (genetic distance and site identity, both using NTS and AAS) gave slightly different results for the *Mixta* sequences. When comparing the results, approximately 38.7% of both the *M. calida* and *M. gaviniae* sequences agreed upon a closest relative. However, the number of sequences closely related to any of the other taxa were slightly different between *M. calida* and *M. gaviniae* (Table 3). There were no genes for which all four analyses agreed on *T. ptyseos* being the closest relative (Table 3). Additionally, 7.5% of the *M. calida* and *M. gaviniae* homologs did not share the same closest relative.

```{r Table3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
CR_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_AllResults_calida.csv", 
                      stringsAsFactors = FALSE)

# table(CR_calida$ALL)
# FALSE  TRUE         # FALSE: 61.3%
# 490    309          # TRUE: 38.7% of M. calida genes had the same closest relative in all four analyses

CR_gaviniae <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_AllResults_gaviniae.csv", 
                        stringsAsFactors = FALSE)

# table(CR_gaviniae$ALL)
# FALSE  TRUE 
# 490    309

all_same_cal <- subset(CR_calida, ALL == TRUE); all_same_gav <- subset(CR_gaviniae, ALL == TRUE)

all_same_species <- data.frame(table(all_same_cal$NT_dist),
                               table(all_same_gav$NT_dist)) %>%
  subset(select = -Var1.1) %>%
  mutate(Var1 = factor(Var1, 
                          levels = Spp, ordered = TRUE)) %>%
  arrange(Var1)

colnames(all_same_species) <- c("Species", "# of M. calida genes", "# of M. gaviniae genes")

all_same_species$Species <- gsub("_", ". ", all_same_species$Species)

kable(all_same_species, 
      caption = "Table 3. The number of *M. calida* and *M. gaviniae* genes wherein all analyses gave the same species as the closest relative.")
```

```{r Table4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
CR_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_AllResults_calida.csv", 
                      stringsAsFactors = FALSE)

CR_gaviniae <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_AllResults_gaviniae.csv", 
                        stringsAsFactors = FALSE)

cal_genus <- subset(CR_calida, Genus != "Nope")
gav_genus <- subset(CR_gaviniae, Genus != "Nope")

all_same_genus <- data.frame(table(cal_genus$Genus),
                             table(gav_genus$Genus))

all_same_genus <- subset(all_same_genus, select = -Var1.1)
colnames(all_same_genus) <- c("Genus", "# of M. calida genes", "# of M. gaviniae genes")

all_same_genus <- all_same_genus %>%
  mutate(Genus = factor(Genus, 
                       levels = c("Pantoea", "Erwinia", "Tatumella", "Enterobacter"), ordered = TRUE)) %>%
  arrange(Genus)

test <- data.frame(Gene = CR_calida$Gene, cal = CR_calida$Genus, gav = CR_gaviniae$Genus)
test$Same <- case_when(test$cal == test$gav ~ test$cal,
                       TRUE ~ "Nah")
test <- subset(test, Same == "Nah")
# table(test$Same)
# FALSE  TRUE 
# 39     760

kable(all_same_genus, 
      caption = "Table 4. The number of *M. calida* and *M. gaviniae* genes wherein all analyses gave the same genus as the closest relative.")
```
